---
title: "Introduction to SAPP"
author: "Mohamed Amine FARES"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to SAPP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4
)
```

## Overview
SAPP (Sector-Adjusted Points Plot) is an R package that visualizes feature dominance in a two-dimensional space. It combines PCA for dimensionality reduction with a novel sector-adjustment mechanism to show which features drive predictions for which observations.

## Quick Start
Load the package and prepare your data:

```{r}
library(SAPP)
data(iris)
X <- iris[, 1:4]
```

Fit a linear model and compute importances:

```{r}
model <- lm(Petal.Width ~ Sepal.Length + Sepal.Width + Petal.Length, data = iris)
imp <- abs(coef(model)[-1])
names(imp) <- c("Sepal.Length", "Sepal.Width", "Petal.Length")
```

Compute per-observation influence and plot:

```{r}
inf <- influence_feature(X, model)
plot_sapp(X, imp, inf, alpha = "auto")
```

## Interpretation
* **Colors:** The feature that most influenced each observation.
* **Point size:** How strongly that feature dominated.
* **Black markers:** Sector centers, based on global importances.
* **Contours:** Where each feature tends to dominate.

## Advanced Usage: Random Forest
SAPP supports tree-based models via SHAP values (requires fastshap):

```{r eval=FALSE}
library(randomForest)
rf <- randomForest(Species ~ ., data = iris, importance = TRUE)
imp_rf <- importance(rf)[, "MeanDecreaseGini"]
names(imp_rf) <- colnames(X)
inf_rf <- influence_feature(X, rf)
plot_sapp(X, imp_rf, inf_rf, alpha = "auto")
```

## References
Lundberg, S. M., and Lee, S.-I. (2017). A Unified Approach to Interpreting Model Predictions. NIPS.
